---
title: "Parenting RCT (Child Analysis)"
author: "Andrew Bogdan"
format: html
editor: visual
---

## Child Analysis

```{r}
#| label: packages
#| message: false

library(tidyverse)
library(haven)
library(knitr)  # For kable, you don't _really_ need it
```

### Data Cleaning

#### Load, Track, Merge

Okay let's take a look at our data.

```{r}
#| label: load-data

DATA_DIR <- "data"

#' Load .sav data from disk and perform some basic universal cleaning.
#' Do not add any input variables to this function. If you make this function
#'  complicated, then when it breaks, you will have to write a simpler one
#'  to load the data with. How annoying would that be?
load_data <- function(sav_filename, column_names) {
  file.path(DATA_DIR, sav_filename) %>%
    read_sav() %>%
    imap(\(column, name) {
      # Remove newlines from the labels so it's easier to read
      attr(column, "label") <- str_replace_all(attr(column, "label"), "\n+", " / ")
      # Save the loaded name for debug purposes
      attr(column, "original_name") <- name
      column
    }) %>%
    as_tibble() %>%
    mutate(index = row_number())
}

# I know double-assignment like this is bad form but I need the long names
#  for code readability and the short names for code debug-ability
english_baseline_df <- load_data("ENGLISH_Child Baseline Data.sav")
english_control_df <- load_data("ENGLISH_Child Follow-Up - Control.sav")
english_treatment_df <- load_data("ENGLISH_Child Follow-Up - Experimental.sav")
spanish_baseline_df <- load_data("SPANISH_Child Baseline Data.sav")
spanish_baseline_2_df <- load_data("SPANISH 2_Child Baseline Data.sav")
spanish_control_df <- load_data("SPANISH_Child Follow-Up - Control.sav")
spanish_treatment_df <- load_data("SPANISH_Child Follow-Up - Experimental.sav")
```

I want usable names now. You'll notice I (above) saved the old names as an attribute in the columns. This is for debugging, as I'm worried that I'll misinterpret the columns or mismatch them columns when merging.

```{r}
#| label: rename-columns

# This document has the name reassignments so this data is a bit more legible.
source("util/column_name_maps.R")

.rename_columns <- function(df, column_name_map) {
  df %>% rename_with(
    .fn = ~ column_name_map[.x],
    .cols = any_of(names(column_name_map))
    )
}

english_baseline_df <- english_baseline_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_english)

english_control_df <- english_control_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_english)

english_treatment_df <- english_treatment_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_english)

spanish_baseline_df <- spanish_baseline_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_spanish)

spanish_baseline_2_df <- spanish_baseline_2_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_spanish)

spanish_control_df <- spanish_control_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_spanish)

spanish_treatment_df <- spanish_treatment_df %>%
  .rename_columns(column_name_map_common) %>%
  .rename_columns(column_name_map_spanish)
```

Before we merge, we want to save some information about where our columns came from. We're going to collect the old column names, new column names, and labels for each source dataframe into a tibble. This is for visual inspection, to make sure we labeled our data correctly.

```{r }
#| label: col-data-tracking
#| eval: false

.dump_column_info <- function(df, postfix = "") {
  tibble(
    column_name = names(df),
    # This is wild, I know
    !!paste0("original_name", postfix) := map_chr(df, ~ attr(.x, "original_name") %||% ""),
    !!paste0("label", postfix) := map_chr(df, ~ attr(.x, "label") %||% "")
  )
}

column_digest <- .dump_column_info(english_baseline_df, "_en_bl") %>%
  full_join(.dump_column_info(english_control_df, "_en_c"), by = "column_name") %>%
  full_join(.dump_column_info(english_treatment_df, "_en_tx"), by = "column_name") %>%
  full_join(.dump_column_info(spanish_baseline_df, "_es_bl"), by = "column_name") %>%
  full_join(.dump_column_info(spanish_baseline_2_df, "_es_bl2"), by = "column_name") %>%
  full_join(.dump_column_info(spanish_control_df, "_es_c"), by = "column_name") %>%
  full_join(.dump_column_info(spanish_treatment_df, "_es_tx"), by = "column_name")

```

Alright now let's merge them. I'd like to make some more accountability on if we are merging the right columns together, but knowing what they look like merged will be super helpful for that.

```{r}
#| label: row-data-tracking
# This cell actually has to run, as the row information we add is used to sort
#  rows.

# Add languages, baseline/control/treatment, legible ID column
.languages = c("english", "spanish")
.group = c("baseline", "control", "treatment")

english_baseline_df <- english_baseline_df %>%
  mutate(
    survey_lang = factor("english", levels = .languages),
    experiment_group = factor("baseline", levels = .group),
    unique_index = paste("en-bl-", index, sep="")
  ) %>%
  relocate(unique_index)

english_control_df <- english_control_df %>%
  mutate(
    survey_lang = factor("english", levels = .languages),
    experiment_group = factor("control", levels = .group),
    unique_index = paste("en-c-", index, sep="")
  ) %>%
  relocate(unique_index)

english_treatment_df <- english_treatment_df %>%
  mutate(
    survey_lang = factor("english", levels = .languages),
    experiment_group = factor("treatment", levels = .group),
    unique_index = paste("en-tx-", index, sep="")
  ) %>%
  relocate(unique_index)

spanish_baseline_df <- spanish_baseline_df %>%
  mutate(
    survey_lang = factor("spanish", levels = .languages),
    experiment_group = factor("baseline", levels = .group),
    unique_index = paste("es-bl-", index, sep="")
  ) %>%
  relocate(unique_index)

spanish_baseline_2_df <- spanish_baseline_2_df %>%
  mutate(
    survey_lang = factor("spanish", levels = .languages),
    experiment_group = factor("baseline", levels = .group),
    unique_index = paste("es-bl2-", index, sep="")
  ) %>%
  relocate(unique_index)

spanish_control_df <- spanish_control_df %>%
  mutate(
    survey_lang = factor("spanish", levels = .languages),
    experiment_group = factor("control", levels = .group),
    unique_index = paste("es-c-", index, sep="")
  ) %>%
  relocate(unique_index)

spanish_treatment_df <- spanish_treatment_df %>%
  mutate(
    survey_lang = factor("spanish", levels = .languages),
    experiment_group = factor("treatment", levels = .group),
    unique_index = paste("es-tx-", index, sep="")
  ) %>%
  relocate(unique_index)
```

This next cell is going to make a bunch of warnings. I am suppressing them against my better judgement— it doesn't like that we're combining columns with haven labels that are different. Some are Spanish, and some are English. I'll remove the labels manually when I get time.

```{r}
#| label: merge
#| warning: false

# This should have 125 columns now.
df <- bind_rows(
  english_baseline_df,
  english_control_df,
  english_treatment_df,
  spanish_baseline_df,
  spanish_baseline_2_df,
  spanish_control_df,
  spanish_treatment_df
)
```

#### Cleaning Specific Columns

**NOTE**: I almost converted the numeric survey responses to numeric with `as.numeric` here. However, that nukes the labels, and I want to preserve those as long as possible. Hence, I'm adopting the custom: convert them to numeric whenever you need them, but don't touch the original data.

We want to match baseline to treatment/control entries. We're going to use phone number to do that; let's clean those up.

```{r}
#| label: clean-phone-number
# This piece of code is a little nasty— I'd love to learn how to
#  make it less so at a different time.

#' This function removes all non-digits and then takes the 
#' last 10 digits to get the phone number without country code.
.clean_phone_number <- function(phone) {
  phone %>%
    na_if("") %>%
    str_replace_all("\\D", "") %>%
    str_replace_all("\\d+(\\d{10})", "\\1")
}

df <- df %>%
  # Apply the clean_phone_number function to every column which starts with
  #  the phrase 'phone_number'
  mutate(across(
    starts_with("phone_number"),
    ~ .clean_phone_number(.x)
    )) %>%
  # We are not going to aggregate across the phone number rows, so we need to
  #  transform row-wise.
  rowwise() %>%
  mutate(
    # List all of the unique phone numbers they submitted. Hopefully, this is
    #  a single-length list for every entry.
    phone_number_unique = list(unique(na.omit(c_across(starts_with("phone_number"))))),
    # The number of phone numbers they submitted; hopefully 1
    phone_number_n_unique = length(phone_number_unique),
    # One of the (hopefully the only) phone number they submitted.
    phone_number = pluck(phone_number_unique, 1, .default=NA)
  ) %>%
  ungroup() %>%
  # List columns really suck so I'm going to kill this one
  select(-phone_number_unique) %>%
  # Count how many entries there are for each phone number
  add_count(phone_number, name="phone_number_n_entries")
  # Reordering makes debugging way easier
  # relocate(unique_index, phone_number, phone_number_n_entries) %>%
  # This is to see if there are duplicate entries
  # arrange(desc(phone_number_n_entries))
```

#### Drop Bad Data

Okay, so we have many (280) submissions which have no phone number associated with them. We also have some phone numbers with up to 4 usages in the same stage of the trial (en-bl-197, 198, 338, and 443). This means we have some additional cleaning to do.

In particular, we need to drop any entries which share both phone number and experimental group. We also have to get rid of treatment/control entries which have no corresponding baseline.

```{r}
#| label: remove-junk

df <- df %>%
  # Do we care about un-finished entries?
  filter(finished == 1) %>%
  # Remove entries which have the same phone number and experimental group
  group_by(phone_number, experiment_group) %>%
  filter(n() == 1) %>%
  ungroup() %>%
  # Let's recount these
  add_count(phone_number, name="num_measurements") %>%
  # Remove control/treatment measurements with no baseline
  filter(experiment_group == "baseline" | num_measurements > 1) %>%
  # Relocate & sort for Andrew reasons
  relocate(unique_index, num_measurements, phone_number) %>%
  arrange(desc(num_measurements), phone_number, experiment_group)

table(df$num_measurements)
```

After this sage of cleaning, we have 778 baseline measurements with no follow-up and 447 baseline/follow-up pairs.

### More Requests

Okay! We have more requests:

-   Aggregate mean responses to each measure
-   Make the 2x long, combining before/after into one row
-   Explain the following missing data:
    -   Adolescent response
    -   Childhood responses
    -   IPPA responses
    -   gender, religion, phone use in follow-ups

First let's take a look at those missing data.

#### (1) Childhood/Adolescent Matching

The percentage of NA in the childhood flourishing measures seems to be exactly the compliment of the NA% of the adolescent flourishing measures, which seems to mean that they were given to different people. Let's check that.

```{r}
#| label: verify-childhood-adolescent-knit
#| eval: false

# I'm sorry that I'm doing a for loop here. This hurts me more than it hurts you.
for (index in 1:12) {
  # The xor (exclusive or) will give TRUE if exactly on of them is NA.
  # This will throw an error if there are any rows with both NAs or both data.
  stopifnot(all(xor(
    is.na(df[[paste0("childhood_", index)]]),
    is.na(df[[paste0("adolescent_", index)]])
  )))
}

```

Okay, so we're good to go. Now this is the perfect use case for coalesce.

```{r}
#| label: coalesce-childhood-adolescent
#| warning: false
# Once again I am killing the 'conflicting labels' warning instead of dealing with it.

# OK but honestly the tidyverse version that the bot tried to give me was so much worse.
for (index in 1:12) {
  df[[paste0("flourishing_", index)]] = coalesce(
    df[[paste0("childhood_", index)]],
    df[[paste0("adolescent_", index)]]
  )
}
```

#### (2) Survey Response Aggregates

We are going to aggregate the following responses:

-   `flourishing` (created from `childhood` and `adolescent`), 12 measurements
-   `ippa` (14 measurements)
-   `patience` (2 measurements)
-   `serenity` (11 measurements)
-   `parent_love` (3 measurements)
-   `child_love` (3 measurements)
-   `extra_binary` (9 measurements): this one is a problem, as #4 is missing in the treatment group and #3 has half a percent missing. It's only for the follow-ups too.
-   `extra_scale` (9 measurements) Also only follow-up.

```{r}
#| label: aggregate

# NOTE: I used matches here so that this line is idempotent. Because it's an
#  average, this makes no difference, but this is better form anyway.
# NOTE2: I used rowwise/summarize/c_across at first but it hung forever.
df <- df %>%
  mutate(across(
    # Is this nasty? I'm sorry
    matches("(flourishing|ippa|patience|serenity|(parent|child)_love|extra_(binary|scale))_\\d+"),
    as.numeric
  )) %>%
  mutate(
    flourishing_mean = rowMeans(select(., matches("flourishing_\\d+")), na.rm = FALSE),
    ippa_mean = rowMeans(select(., matches("ippa_\\d+")), na.rm = FALSE),
    patience_mean = rowMeans(select(., matches("patience_\\d+")), na.rm = FALSE),
    serenity_mean = rowMeans(select(., matches("serenity_\\d+")), na.rm = FALSE),
    parent_love_mean = rowMeans(select(., matches("parent_love_\\d+")), na.rm = FALSE),
    child_love_mean = rowMeans(select(., matches("child_love_\\d+")), na.rm = FALSE),
    # flourishing_mean = rowMeans(select(., matches("extra_binary")), na.rm = FALSE),
    extra_scale_mean = rowMeans(select(., matches("extra_scale_\\d+")), na.rm = FALSE),
  )
```

### Output

This is the end of this document: output artifacts.

#### Column Digest

This creates a table where each row is one column in our dataframe. It collects information about every column so you can see what's going on. It's for documentation and debugging.

This relies on you running the @ref(chunk:col-data-tracking) chunk.

```{r}
#| label: write-column-digest
#| eval: false

DOCS_DIR = "tables"

#' I make a function and then call it exactly once— why?
#'  I actually use this function a bunch while debugging. I'd move it to a
#'  separate file but that would be confusing.
.make_dtypes <- function(df) {
  df %>%
    summarise(across(everything(), ~ class(.x)[1])) %>%
    pivot_longer(everything(), names_to = "column_name", values_to = "dtype")
}

dtypes <- .make_dtypes(df)

# This gets the percent of entries that are NA based on what experimental
#  group the entry is part of.
# I'd like to make this a function to, but I don't have an idiomatic way to 
#  abstract the grouping without making it mandatory. Hence, until I need it,
#  it won't get the dev time.
percent_na <- df %>%
  # Fill any empty strings with NA
  mutate(across(
    where(is_character),
    ~ na_if(.x, "")
    )) %>%
  group_by(experiment_group) %>%
  # Get the percent of each group which is NA
  summarise(across(everything(), ~ mean(is.na(.x)))) %>%
  # These group names will become column names, so let's be clearer
  mutate(experiment_group = paste("percent_na", experiment_group, sep="_")) %>%
  # This rotates the data so I can scroll down instead of sideways
  pivot_longer(
    cols = -experiment_group,
    names_to = "column_name",
    values_to = "percent_na"
  ) %>%
  # This turns experiment_group and percent_na columns into three columns,
  #  baseline_percent_na, control_percent_na, treatment_percent_na;
  #  those names are stored in the experiment_group column right now.
  pivot_wider(
    names_from = experiment_group,
    values_from = "percent_na"
  )

# Now let's add this info to the column digest and sort it
column_digest <- column_digest %>%
  full_join(percent_na, by = "column_name") %>%
  full_join(dtypes, by = "column_name") %>%
  # We do these in reverse order
  relocate(starts_with("label")) %>%
  relocate(starts_with("original_name")) %>%
  relocate(starts_with("percent_na")) %>%
  relocate(column_name, "dtype")

# Let's save this as a Markdown artifact
column_digest %>%
  kable(format = "markdown") %>%
  writeLines(con = file.path(DOCS_DIR, "column_digest.md"))

# And also as a csv so I can bring it into google docs
column_digest %>%
  write_csv("debug/column_digest.csv")
```
