---
title: "Parenting RCT (Child Analysis)"
author: "Andrew Bogdan"
format: html
editor: visual
---

## Child Analysis

```{r message=FALSE}
#| label: packages
# TODO: Pin versions in knitted output for reproducability
library(tidyverse)
library(haven)
```

Okay let's take a look at our data.

```{r}
#| label: load-data

DATA_DIR <- "data"

#' Load .sav data from disk and perform some basic universal cleaning.
#' Do not add any input variables to this function. If you make this function
#'  complicated, then when it breaks, you will have to write a simpler one
#'  to load the data with. How annoying would that be?
load_data <- function(sav_filename, column_names) {
  file.path(DATA_DIR, sav_filename) %>%
    read_sav() %>%
    imap(\(column, name) {
      # Remove newlines from the labels so it's easier to read
      attr(column, "label") <- str_replace_all(attr(column, "label"), "\n+", " / ")
      # Save the loaded name for debug purposes
      attr(column, "original_name") <- name
      column
    }) %>%
    as_tibble()
}

# I know double-assignment like this is bad form but I need the long names
#  for code readability and the short names for code debug-ability
english_baseline_df <- load_data("ENGLISH_Child Baseline Data.sav")
english_control_df <- load_data("ENGLISH_Child Follow-Up - Control.sav")
english_treatment_df <- load_data("ENGLISH_Child Follow-Up - Experimental.sav")
spanish_baseline_df <- load_data("SPANISH_Child Baseline Data.sav")
spanish_baseline_2_df <- load_data("SPANISH 2_Child Baseline Data.sav")
spanish_control_df <- load_data("SPANISH_Child Follow-Up - Control.sav")
spanish_treatment_df <- load_data("SPANISH_Child Follow-Up - Experimental.sav")
```

I want usable names now. You'll notice I (above) saved the old names as an
attribute in the columns. This is for debugging, as I'm worried that I'll 
misinterpret the columns or mismatch them columns when merging.
```{r}
#| label: rename-columns

# This document has the name reassignments so this data is a bit more legible.
source("util/column_name_maps.R")

rename_columns <- function(df, column_name_map) {
  df %>% rename_with(
    .fn = ~ column_name_map[.x],
    .cols = any_of(names(column_name_map))
    )
}

english_baseline_df <- english_baseline_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_english)

english_control_df <- english_control_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_english)

english_treatment_df <- english_treatment_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_english)

spanish_baseline_df <- spanish_baseline_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_spanish)

spanish_baseline_2_df <- spanish_baseline_2_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_spanish)

spanish_control_df <- spanish_control_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_spanish)

spanish_treatment_df <- spanish_treatment_df %>%
  rename_columns(column_name_map_common) %>%
  rename_columns(column_name_map_spanish)
```

This here is a cell which will dump to a CSV the old column names, new column
names, and label of the dataframes. This is for visual inspection, to make 
sure we labeled our data correctly.
```{r}
#| label: dump-column-debug
#| eval: false

DEBUG_COLUMN_NAMES_DIR <- "debug_column_names"

.debug_dump_column_info_to_csv <- function(df, csv_filename) {
  tibble(
    name = names(df),
    original_name = map_chr(df, ~ attr(.x, "original_name")),
    label = map_chr(df, ~ attr(.x, "label"))
  ) %>% write_csv(file.path(DEBUG_COLUMN_NAMES_DIR, csv_filename))
}

english_baseline_df %>% .debug_dump_column_info_to_csv("english_baseline.csv")
english_control_df %>% .debug_dump_column_info_to_csv("english_control.csv")
english_treatment_df %>% .debug_dump_column_info_to_csv("english_treatment.csv")
spanish_baseline_df %>% .debug_dump_column_info_to_csv("spanish_baseline.csv")
spanish_baseline_2_df %>% .debug_dump_column_info_to_csv("spanish_baseline_2.csv")
spanish_control_df %>% .debug_dump_column_info_to_csv("spanish_control.csv")
spanish_treatment_df %>% .debug_dump_column_info_to_csv("spanish_treatment.csv")
```

